---
title: "Smart Data: Key Features"
output: rmarkdown::html_vignette
vignette: >
 %\\VignetteIndexEntry{Smart Data: Key Features}
 %\\VignetteIndexEngine{knitr::rmarkdown}
 %\\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup 

To demonstrate the key features of the `smart.data` package, we'll use the `mtcars` data set:

```{r setup}
library(smart.data)

smrt <- smart.data$new(x = mtcars, name = "smart_cars")
```

# Naming API

```{r NAMING_RULE}
# `make_model = "rn"` is needed because row names are retained when the 
# 	input data is converted to a `data.table` object
smrt$naming.rule(
	!!!c(make_model = "rn"
			 , (\(x) rlang::set_names(x, toupper(x)))(names(mtcars))))$
	enforce.rules(for_naming)

print(head(smrt$data))
```

# Taxonomy API

## Defining Taxonomy Terms

Next, we'll define a set of taxonomy terms and map them to subsets of the field names.  

> *Note*: if an pending naming map exists, it will automatically be enforced to ensure taxonomy field selection is based on current data names:

Taxonomy terms:

- **identifier**: make_model
- **performance**: MPG, QSEC, HP
- **metrics**: CYL, DISP, DRAT, WG, GEAR, CARB
- **style**: VS, AM

```{r TAXONOMY_DEF}
identifier <- taxonomy(term = "identifier", desc = "Make and Model"
											 , fields = c("make_model"))
performance <- taxonomy(term = "performance", desc = "Performance stats"
												, fields = c("MPG", "QSEC", "HP"))
metrics <- taxonomy(term = "metrics", desc = "Physical Metrics"
										, fields = c("CYL", "DISP", "DRAT", "WT", "GEAR", "CARB"))
style <- taxonomy(term = "characteristics", desc = "Categorical Descriptors"
									, fields = c("VS", "AM"))

smrt$
	taxonomy.rule(identifier, performance, metrics, style)$
	enforce.rules(for_usage)
```

## Using Taxonomy Terms 

To use the defined terms, class method `$use()` is invoked.  The output can be tailored depending on the arguments supplied:

```{r TAXONOMY_USE}
# Use everything
smrt$use() |> head()

# Use 'identifier' fields
smrt$use(identifier) |> head()

# Use 'performance' fields
smrt$use(performance) |> head()

# Use 'metrics' fields omitting field "DISP"
smrt$use(metrics, omit=DISP) |> head()

# Use 'style' and 'identifier' fields also retaining field "QSEC
smrt$use(style, identifier, retain = QSEC) |> head()

# Use 'style' and 'identifier' fields also retaining fields "DISP" and "QSEC"
smrt$use(style, identifier, retain = c(DISP, QSEC)) |> head()

# Use 'style' and 'identifier' fields also retaining fields "DISP" and "QSEC" 
#		filtering for rows where field 'make_model' contains the pattern "Merc"
smrt$use(style, identifier, retain = c(DISP, QSEC)
				 , subset = grepl("Merc", make_model)) |> head()
```

# Using the Smart Cache 

Each `smart.data` object can be given an name during class invocation or by assigning directly after the fact.  When names are compliant for use with `cachem`, the class instance can be sent to the cache contained within the package environment:

```{r SMART_CACHE}
smrt$cache_mgr(action = register)

# View existing cached objects
get.smart(list.only = TRUE)

# Verify equivalent calls
identical(get.smart(smart_cars)$data, smrt$data)

# Use cached objects with taxonomy
get.smart(smart_cars)$
	use(performance, identifier, subset = grepl("Merc", make_model))
```

If the underlying data is changed, the cached instance can be updated by using `$cache_mgr(action = replace)`.

The benefit of caching is that the cached instance can be invoked from anywhere in your workflow without worrying about environment scope (except for parallel computing contexts).
